---
title: "Deep Learning Customer Retention with ANN"
author: "Antonis Angelakis"
date: "August 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Import Libraries

```{r Import libraries,warning=FALSE,message=FALSE}
library(plyr)
library(dplyr)
library(ggpubr)
library(car)
library(caTools)
library(h2o) 
```
**h20 install  --> install.packages("h2o", type="source",repos=(c("http://h2o-release.s3.amazonaws.com/h2o/latest_stable_R"))) worked for my system **

###Import Dataset

```{r}
Customer_Churn <- read.csv("../input/WA_Fn-UseC_-Telco-Customer-Churn.csv")
```

#Data pre-processing

```{r Data pre-processing}
##Delete CustomerID
Customer_Churn = subset(Customer_Churn, select = -c(customerID))

##Missing Values
na_count <-sapply(Customer_Churn, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count) 
na_count

##Find NA observations
Customer_Churn_NA <- subset(Customer_Churn, is.na(Customer_Churn$TotalCharges))

##Remove them 
Customer_Churn <- na.omit(Customer_Churn)

##Change SeniorCitizen (0,1) to factors
Customer_Churn$SeniorCitizen <- as.factor(
  mapvalues(Customer_Churn$SeniorCitizen,
            from=c("0","1"),
            to=c("No", "Yes"))
)

##Change tenure to numerical
Customer_Churn[5] <- lapply(Customer_Churn[5], as.numeric)
```
**CustomerID is useless for our model. We have 11 NA observations for the column TotalCharges.**
**We delete them. We change SeniorCitizen to No/Yes and transform tenure from int to num. **

##Explore Variables
```{r explore, echo = TRUE}
str(Customer_Churn)
summary(Customer_Churn)
```

##Plot Categorical Variables
```{r Plot Categorical Variables}
plot(Customer_Churn$gender, main="gender")
plot(Customer_Churn$SeniorCitizen, main="SeniorCitizen")
plot(Customer_Churn$Partner, main="Partner")
plot(Customer_Churn$Dependents, main="Dependents")
plot(Customer_Churn$PhoneService, main="PhoneService")
plot(Customer_Churn$MultipleLines, main="MultipleLines")
plot(Customer_Churn$InternetService, main="InternetService")
plot(Customer_Churn$OnlineSecurity, main="OnlineSecurity")
plot(Customer_Churn$OnlineBackup, main="OnlineBackup")
plot(Customer_Churn$DeviceProtection, main="DeviceProtection")
plot(Customer_Churn$TechSupport, main="TechSupport")
plot(Customer_Churn$StreamingTV, main="StreamingTV")
plot(Customer_Churn$StreamingMovies, main="StreamingMovies")
plot(Customer_Churn$Contract, main="Contract")
plot(Customer_Churn$PaperlessBilling, main="PaperlessBilling" )
plot(Customer_Churn$PaymentMethod, main="PaymentMethod" )
plot(Customer_Churn$Churn,  main="Churn")
```

##Artificial Neural Network 
###Encoding Categorical Variables as factors
**h2o library needs categorical variables to be converted into numerical.**
```{r Encoding the categorical variables as factors}
Customer_Churn$gender = as.numeric(factor(Customer_Churn$gender,
                                      levels = c('Female', 'Male'),
                                      labels = c(1, 2)))
Customer_Churn$SeniorCitizen = as.numeric(factor(Customer_Churn$SeniorCitizen,
                                          levels = c('No', 'Yes'),
                                          labels = c(1, 2)))
Customer_Churn$Partner = as.numeric(factor(Customer_Churn$Partner,
                                          levels = c('No', 'Yes'),
                                          labels = c(1, 2)))
Customer_Churn$Dependents = as.numeric(factor(Customer_Churn$Dependents,
                                          levels = c('No', 'Yes'),
                                          labels = c(1, 2)))
Customer_Churn$PhoneService = as.numeric(factor(Customer_Churn$PhoneService,
                                              levels = c('No', 'Yes'),
                                              labels = c(1, 2)))
Customer_Churn$MultipleLines = as.numeric(factor(Customer_Churn$MultipleLines,
                                                levels = c('No', 'No phone service','Yes'),
                                                labels = c(1, 2,3)))
Customer_Churn$InternetService = as.numeric(factor(Customer_Churn$InternetService,
                                                 levels = c('DSL', 'Fiber optic','No'),
                                                 labels = c(1, 2,3)))
Customer_Churn$OnlineSecurity = as.numeric(factor(Customer_Churn$OnlineSecurity,
                                                   levels = c('No', 'No internet service','Yes'),
                                                   labels = c(1, 2,3)))
Customer_Churn$OnlineBackup = as.numeric(factor(Customer_Churn$OnlineBackup,
                                                  levels = c('No', 'No internet service','Yes'),
                                                  labels = c(1, 2,3)))
Customer_Churn$DeviceProtection = as.numeric(factor(Customer_Churn$DeviceProtection,
                                                  levels = c('No', 'No internet service','Yes'),
                                                  labels = c(1, 2,3)))
Customer_Churn$TechSupport = as.numeric(factor(Customer_Churn$TechSupport,
                                                  levels = c('No', 'No internet service','Yes'),
                                                  labels = c(1, 2,3)))
Customer_Churn$StreamingTV = as.numeric(factor(Customer_Churn$StreamingTV,
                                                  levels = c('No', 'No internet service','Yes'),
                                                  labels = c(1, 2,3)))
Customer_Churn$StreamingMovies = as.numeric(factor(Customer_Churn$StreamingMovies,
                                                  levels = c('No', 'No internet service','Yes'),
                                                  labels = c(1, 2,3)))
Customer_Churn$Contract = as.numeric(factor(Customer_Churn$Contract,
                                                  levels = c('Month-to-month', 'One year','Two year'),
                                                  labels = c(1, 2,3)))
Customer_Churn$PaperlessBilling = as.numeric(factor(Customer_Churn$PaperlessBilling,
                                            levels = c('No', 'Yes'),
                                            labels = c(1, 2)))
Customer_Churn$PaymentMethod = as.numeric(factor(Customer_Churn$PaymentMethod,
                                            levels = c('Electronic check', 'Mailed check','Credit card (automatic)','Bank transfer (automatic)'),
                                            labels = c(1, 2,3,4)))
```

###Recode Churn from 1 or 2 to 0 or 1
```{r Recode Churn from (1-2) to (0-1)}
Customer_Churn$Churn = as.numeric(factor(Customer_Churn$Churn,
                                                 levels = c('No', 'Yes'),
                                                 labels = c(1, 2)))
Customer_Churn$Churn<-recode(Customer_Churn$Churn,"1=0; 2=1")
```
**Neural network works only if our dependent variable is 0 or 1. Values must be near 0**

##Splitting the dataset into training and test set
```{r Splitting the dataset into training and test set}
set.seed(123)
split = sample.split(Customer_Churn$Churn, SplitRatio = 0.8)
training_set = subset(Customer_Churn, split == TRUE)
test_set = subset(Customer_Churn, split == FALSE)
```

##Feature Scaling
```{r Feature Scaling}
training_set[-20] = scale(training_set[-20])
test_set[-20] = scale(test_set[-20])
```
**For ANN to work we need to scale all our independent variables.**

##Fit ANN to training set
```{r Fit ANN to training set,message=FALSE,results = 'hide'}
h2o.init(nthreads = -1)
model = h2o.deeplearning(y = 'Churn',
                         training_frame = as.h2o(training_set),
                         activation = 'Rectifier',
                         hidden = c(3,6),
                         epochs = 1000,
                         train_samples_per_iteration = -2)
```
**After too much testing, most stable prediction results came for 3 hidden layers with 6 neurons.There are 19 neurons in the input and 1** 
**in the ouput layer.Too many neurons in the hidden layers may result in overfitting. Number of epochs is 500.**

##Predict test set results & Confusion Matrix
```{r Predict test set results & Confusion Matrix, message = FALSE}
y_pred = h2o.predict(model, newdata = as.h2o(test_set[-20]))
y_pred = (y_pred > 0.5)
y_pred = as.vector(y_pred)
cm = table(test_set[, 20], y_pred)
cm
accuracy <- (cm[1,1]+cm[2,2]) / 1407 
specificity <- cm[1,1]/ 1033
precision <- cm[2,2]/(cm[1,2]+cm[2,2])
print("Accuracy (Overall, how often is the classifier correct?) ")
accuracy 
print("Specificity (When it's actually no, how often does it predict no?) ")
specificity  
print("Precision (When it predicts yes, how often is it correct?) ")
precision 
h2o.shutdown(prompt = FALSE)
```
